The resurgence of the linear optics interferometer --- recent advances \& applications

### 2018-01-03 paper.tex
Revised QKD paragraph in Section 10.1.

Linear optics play a prominent role in quantum information processing. Photons make fantastic `flying' qubits, and are readily used for quantum communication \cite{bib:Northup14} and quantum key distribution \cite{bib:Lo14}. In 2001, Knill, Laflamme and Milburn (KLM) showed that efficient quantum computing is possible using only linear optical components, that is single photons, beamsplitters, phase shifters and photon counting \cite{bib:KLM01}. Furthermore, an optical implementation of quantum protocols can potentially piggyback on existing communication infrastructure. Combined with the relative ease of transporting quantum states of light, as compared to say trapped ions or superconducting qubits, linear optics forms a powerful platform for quantum information processing.

@@

Better single-photon sources (see Section \ref{sec:singlephotons}) are part of the solution. Improvement in control of their degrees of freedom have also led to more efficient designs of circuits \cite{bib:Zhou11,bib:Lanyon09} that enabled the optical implementation of quantum teleportation of the state of a single photon \cite{bib:Wang15}. In this way, three-qubit \cite{bib:Lanyon09,bib:Micuda13,bib:Patel16} and four-qubit \cite{bib:Starek16} gates have also been implemented using LOQC. \sihui{We also briefly mention here that single photons and waveplates are used in a related application, namely, quantum key distribution (QKD). While QKD fundamentally does not require interferometry \cite{bib:BennetBrassard84,bib:Ekert91}, some QKD protocols rely on the Mach-Zehnder \cite{Bennett92, bib:Hughes2000,bib:Sasaki14,bib:Takesue15} and Franson interferometers \cite{bib:Ali-Khan07,bib:Brougham13}. Practical quantum key distribution relying on photonic technology is reviewed in \cite{bib:Scarani09,bib:Hoi-Kwong14}.
}

@@

  Although there has been tremendous progress in linear optics, it seems likely that an efficient, and fault-tolerant universal quantum computer will require a hybrid system. Nonetheless, linear optics remains fascinating platform for applications and foundation work.

@@

where ${\textbf Per}()$ denotes the permanent of a matrix, $S=\{s_1,\dots,s_m\}$ denotes an output photon-number configuration, and $U_S$ is obtained from the interferometer unitary $U$ by keeping its first n columns and repeating its $j$th row $s_j$ times. The appearance of the permanent, which is \#P-hard in general (a complexity class even harder than NP-hard), in this statistics contributes to the hardness of the boson sampling problem. Note that the number of configurations $|S|=\binom{n+m-1}{n}$ in the output superposition scales exponentially with $m$. For this reason, the boson sampler does not let us \emph{calculate} matrix permanents, as this would require knowing individual amplitudes with high precision, which would require an exponential number of samples. Rather, the boson sampler samples across a distribution of permanents without actually revealing any of them to the experimenter. Nonetheless, this is a classically intractable problem. It is also known that boson sampling is as hard as stimulating the short time evolution of a certain Hamiltonian\cite{bib:Peropadre17}.
%>>>>>>> origin/master:(PR final edits) linear_optics_review/paper.tex

@@

\peter{
Some of the protocols thus far require switching between different optical modes. This is achieved experimentally via a Mach-Zehnder interferometer with a dynamic, classically-controlled phase-shift in one path. The variable phase may be implemented in a number of ways using various types of phase modulators, such as electro-optic or acousto-optic modulators. With this construction setting the phase to $\phi=0$ performs the identity operation, whereas for $\phi=\pi$ it performs a swap gate. This is shown in Fig. \ref{figMZ}.
}

>>>>

Linear optics play a prominent role in quantum information processing. Photons make fantastic `flying' qubits, and are readily used for quantum communication \cite{bib:Northup14} and quantum key distribution \cite{bib:Lo14}. In 2001, Knill, Laflamme and Milburn (KLM) showed that efficient quantum computing is possible using only linear optical components, that is single photons, beamsplitters, phase shifters and photon counting  \sihui{for spatially encoded qubits \cite{bib:KLM01}. At about the same time, Koashi, Yamamoto, and Imoto essentially came to the same result for polarization-encoded qubits \cite{Koashi01}.} Furthermore, an optical implementation of quantum protocols can potentially piggyback on existing communication infrastructure. Combined with the relative ease of transporting quantum states of light, as compared to say trapped ions or superconducting qubits, linear optics forms a powerful platform for quantum information processing.

@@

Better single-photon sources (see Section \ref{sec:singlephotons}) are part of the solution. Improvement in control of their degrees of freedom have also led to more efficient designs of circuits \cite{bib:Zhou11,bib:Lanyon09} that enabled the optical implementation of quantum teleportation of the state of a single photon \cite{bib:Wang15}. In this way, three-qubit \cite{bib:Lanyon09,bib:Micuda13,bib:Patel16} and four-qubit \cite{bib:Starek16} gates have also been implemented using LOQC.

@@

  Although there has been tremendous progress in linear optics, it seems likely that an efficient, and fault-tolerant universal quantum computer will require a hybrid system. \sihui{We also briefly mention here that single photons and waveplates are used in a related application, namely, quantum key distribution (QKD). While QKD fundamentally does not require interferometry \cite{bib:BennetBrassard84,bib:Ekert91}, some QKD protocols rely on the Mach-Zehnder \cite{Bennett92, bib:Hughes2000,bib:Sasaki14,bib:Takesue15} and Franson interferometers \cite{bib:Ali-Khan07,bib:Brougham13}. Practical quantum key distribution relying on photonic technology is reviewed in \cite{bib:Scarani09,bib:Hoi-Kwong14}, and incredibly has been implemented at a distance of 1200 km by Earth-satellite transmissions \cite{bib:Liao17}. On the other hand, optimal quantum-optical cloning implementations \cite{bib:Fan2014} based on Mach-Zehnder interferometers have been applied to perform attacks on known quantum cryptographic schemes \cite{bib:Bartkiewicz2013,bib:Bartkiewicz2017}. Thus, linear optics remains fascinating platform for applications and foundation work. 
}

@@

%>>>>>>> origin/master:(PR final edits) linear_optics_review/paper.tex

@@

\peter{
Some of the protocols thus far require switching between different optical modes. \sihui{One way to achieve this experimentally for time-bin qubits is} via a Mach-Zehnder interferometer with a dynamic, classically-controlled phase-shift in one path. The variable phase may be implemented in a number of ways using various types of phase modulators, such as electro-optic or acousto-optic modulators. With this construction setting the phase to $\phi=0$ performs the identity operation, whereas for $\phi=\pi$ it performs a swap gate. This is shown in Fig. \ref{figMZ}.
}
